Spfy: a graph-based genome sequence database for real-time prediction of \textit{E. coli} phenotypes

### 2018-01-16 paper-webserver.tex
EDIT: some things from proposal

Spfy, uses graph data structures to store and retrieve results for computational workflows, facilitating the management and querying of tens of thousands of whole-genome \textit{E. coli} sequences, and efficient downstream processing.
Current comparative computational workflows chain different analysis software, but lack storage and retrieval methods for generated results.
By making the storage and retrieval of results part of the platform, with data effectively linked to the organisms of interest through a standardized ontology, we can mitigate the recomputing of analyses.
Within Spfy, the output from all analyses is stored, and linked together in the context of a genome graph. This graph also stores metadata for each genome, facilitating inquiries ranging from population genomics to epidemiological investigations.
Integrated data storage will be necessary as publicly available whole genome sequencing data for bacterial pathogens currently numbers in the tens of thousands, with hundreds of thousands set to be available within the next few years. \par

@@

% para: semantic web in spfy
In Spfy, the results from all the analysis modules are converted into graph objects, which are used to update the main graph database. These graph objects, which consist of nodes and edges, must be annotated with common terminology. Spfy uses annotations from the GenEpiO \citep{griffiths2017context}, FALDO \citep{bolleman2016faldo}, and TypOn \citep{vaz2014typon} ontologies, as well as custom terminology to enable graph traversals and related data points to be inferred from the ontology (see Figure \ref{fig-ontology}). The use of semantic web technologies allows results to be linked to the genomes they were computed from, and for all data points to share a common data structure that can be queried together.

@@

Like IRIDA and BAP, Spfy automates workflows for users, and like Galaxy, Spfy uses task queues to distribute selected analysis. File uploads begin through the ReactJS-based website, where user-defined analyses options are selected. To these concepts we add Docker containerization for task queue workers, allowing analysis software to safely run in parallel.
% re: Matt "2. When comparing to IRIDA, BAP etc., can mention some differences with Superphy, e.g. the storage of interim result data that allows downstream integrated analysis"
For result storage, existing workflow technologies use relational tables \cite{goecks2010galaxy}, or store resulting files to disk \cite{thomsen2016bacterial}.
Because output from these programs is user-specific or transitory, results from identical comparisons are often recomputed. Additionally, output from different analyses are structured using distinct terminology and formats, which must be converted before they can be compared. Without a unified structure, these conversions quickly become impractical for broad usage. Graph-based storage of all results solves these problems.
% as before
To avoid proliferating ontologies, and to allow Spfy to integrate with existing ones, annotations from the GenEpiO \cite{griffiths2017context}, FALDO \cite{bolleman2016faldo}, and TypOn \cite{vaz2014typon} ontologies are used to describe biological data.
The entire platform is packaged using Docker-Compose, and can be recreated with a simple command. \par

>>>>

Current comparative computational workflows chain different analysis software, but lack storage and retrieval methods for generated results.
Spfy uses graph data structures to store and retrieve results for computational workflows, facilitating the management and querying of tens of thousands of whole-genome \textit{E. coli} sequences, and efficient downstream processing.
By making the storage and retrieval of results part of the platform, with data effectively linked to the organisms of interest through a standardized ontology, we can mitigate the recomputing of analyses.
Within Spfy, the output from all analyses is stored, and linked together in the context of a genome graph. This graph also stores metadata for each genome, facilitating inquiries ranging from population genomics to epidemiological investigations.
Integrated data storage will be necessary as publicly available whole genome sequencing data for bacterial pathogens currently numbers in the tens of thousands, with hundreds of thousands set to be available within the next few years. \par

@@

% para: semantic web in spfy
In Spfy, the results from all the analysis modules are converted into graph objects, which are used to update the main graph database.
These graph objects, which consist of nodes and edges, must be annotated with common terminology.
To avoid proliferating ontologies, and to allow Spfy to integrate with existing ones, annotations from the GenEpiO \cite{griffiths2017context}, FALDO \cite{bolleman2016faldo}, and TypOn \cite{vaz2014typon} ontologies are used to describe biological data.
The use of semantic web technologies allows results to be linked to the genomes they were computed from, and for all data points to share a common data structure that can be queried together.

@@

Like IRIDA and BAP, Spfy automates workflows for users, and like Galaxy, Spfy uses task queues to distribute selected analysis. File uploads begin through the ReactJS-based website, where user-defined analyses options are selected. To these concepts we add Docker containerization for task queue workers, allowing analysis software to safely run in parallel.
% re: Matt "2. When comparing to IRIDA, BAP etc., can mention some differences with Superphy, e.g. the storage of interim result data that allows downstream integrated analysis"
For result storage, existing workflow technologies use relational tables \cite{goecks2010galaxy}, or store resulting files to disk \cite{thomsen2016bacterial}.
Because output from these programs is user-specific or transitory, results from identical comparisons are often recomputed. Additionally, output from different analyses are structured using distinct terminology and formats, which must be converted before they can be compared. Without a unified structure, these conversions quickly become impractical for broad usage. Graph-based storage of all results solves these problems.
% as before
The entire platform is packaged using Docker-Compose, and can be recreated with a simple command. \par
